### 2015-06-14 alubp.tex
Refining the meaning of DP, making the intro more precise
ALUBP cleaned

@@

Given $J\in \R^n$, its approximate least upper bound (ALUB) projection is denoted by $\hg J$ and is defined as 

@@

(\hg J)(i)\stackrel{\Delta}{=}(\Phi r_{e_i,J})(i), \mb i=1,\ldots,n,
where $r_{e_j}$ is a solution to the linear program in \eqref{alubplp} for $c=e_j$, and $e_j$ is the same as in Definition~\ref{lubpop}.
r_{c,J}\stackrel{\Delta}{=} & \underset{r\in \N}{\argmin}\mb c^\top \Phi r,\nn\\
& \text{s.t.}\,\,\, W^\top E \Phi r\geq W^\top HJ,\,\, W \in \R^{nd\times m}_+ .

@@

Note that $W$ in \eqref{alubplp} is the same matrix that is used in \eqref{grlp} and satisfies Assumption~\ref{wassump}.

@@

For $J_1, J_2\in \R^n$ such that $J_1\geq J_2$, we have $\hg J_1\geq \hg J_2$.

@@

The proof follows from Assumptions~\ref{wassump} and ~\ref{one} using arguments along the lines of Lemma~\ref{gmonotone}.

@@

Let $J_1\in \R^n$ and $t\in \R$ be a constant. If $J_2=J_1+t\one$, then $\hg J_2=\hg J_1+\alpha t\one$.

@@

The proof follows from Assumption~\ref{wassump} and ~\ref{one}, as well as Lemma~\ref{lpsol} using arguments along the lines of Lemma~\ref{gshift}. In particular, consider the $i^{th}$ linear program corresponding to $\hg J_1$ and $\hg J_2$. Now, the result follows by letting $A=W^\top E \Phi$, $b=W^\top H J$, $c=e_i$, $b_0=\alpha t \mathbf{1}$, $x_0=\alpha t e_i$.

@@

@@

The operator $\hg \colon \R^n\ra \R^n$ obeys the $\max$-norm contraction property with factor $\alpha$ and the following iterative scheme based on the ALUB projection operator $\hg$, see \eqref{apvi}, converges to a unique fixed point $\hv$.

@@

@@

Follows along similar lines as the proof of Theorem~\ref{gmaxcontra}.

@@

@@

The unique fixed point $\hv$ of the iteration in \eqref{apvi} and the solution $\hj$ of the GRLP \eqref{grlp} obey $\hj\geq\hv$.

@@

I think the proof would be much easier to understand if we introduced $r_{c,J} = \argmin_{r} \{ c^\top \Phi r \,|\, W^\top E \Phi r \ge W^\top H J\}$. We could start by saying we need $V_1\le V_0 \le \hj$ because then Lemma~\ref{tgmonotone} will give the desired result. Next, $V_0\le \hj$ holds because on the one hand, $\hj(i) \ge (\Phi r_i)(i)$ for any $i$, while on the other hand,

@@

@@

Let $\hv$ be the fixed point of the iterative scheme in \eqref{apvi} and let $\bj$ be the best possible LUB approximation of $J^*$ as in Definition~\ref{bestproj}. Then,

@@

||J^*-\hv||_\infty\leq \frac{||J^*-\bj||_\infty+||\Gamma J^*-\hg J^*||_\infty}{1-\alpha}.

@@

Let $\epsilon=||J^*-\bj||_\infty$, and $\{V_n\},n\geq 0$ be the iterates of the scheme in \eqref{apvi} with $V_0=\hg J^*$, then
\todoc{Can the proof be simplified similarly to that of Theorem~\ref{fxpres}? Not a high priority.}
||J^*-\hg J^*||_\infty&\leq||J^*-\Gamma J^*||_\infty+||\Gamma J^*-\hg J^*||_\infty\nn\\
&= \epsilon+\beta,
where $\beta=||\Gamma J^*-\hg J^*||_\infty$. Now

@@

%&=\frac{\epsilon+\beta}{1-\alpha}.

@@

@@

@@

Let $\hv$, $\bj$ be as in Theorem~\ref{mt1} and let $r^*\stackrel{\Delta}{=}\argmin_{r\in \R^k}||J^*-\Phi r||_\infty$, then

@@

The result is obtained by using Lemma~\ref{bestbnd} to replace the term $||J^*-\bj||_\infty$ in Theorem~\ref{mt1}.

>>>>

Recall that $W\in \R^{nd\times m}$ is the matrix in \eqref{grlp} which is assumed to satisfy Assumption~\ref{wassump}.

@@

Given $J\in \R^n$ and the nonnegative valued vector $c\in \R^n_+$, define $r_{c,J}'$ to be the solution to 
 \underset{r\in \R^k}{\min}& \,\mb c^\top \Phi r\,,\nn\\
 \text{s.t.}& \,\,\, W^\top E \Phi r\geq W^\top HJ\,,\qquad r \in \N\,.
Then, for $J\in \R^n$, $\hg J$, the approximate least upper bound (ALUB) projection of $J$ is defined as 

@@

(\hg J)(i)\eqdef(\Phi \har_{e_i,J})(i), \mb i=1,\ldots,n\,.

@@

\noindent 
Note that under our assumptions on $\N$, $\hg$ is well-defined.
We now show that $\hg$ satisfies the monotone property and is linear along $[\one]$-rays with a factor smaller than one and hence it will follow that it is a contraction.

@@

For $J_1, J_2\in \R^n$ such that $J_1\leq J_2$, we have $\hg J_1\leq \hg J_2$.

@@

By \cref{wassump}, $W^\top H: \R^n \to \R^m$ is monotone.
The proof then follows along the lines of Lemma~\ref{gmonotone}.

@@

Let $J\in \R^n$ and $t\in \R$ be a constant. Then $\hg (J+t\one)=\hg J+\alpha t\one$.

@@

The statement follows from Assumptions \ref{one}, \ref{wassump} and \ref{ass:n}, 
as well as Lemma~\ref{lpsol} using arguments along the lines of Lemma~\ref{gshift}. 
In particular, consider the $i^{th}$ linear program corresponding to $\hg J$ and $\hg (J+t\one)$. 
Now, the result follows by letting $A=W^\top E \Phi$, $b=W^\top H J$, 
$c=e_i$, $b_0=\alpha t \mathbf{1}$, 
$x_0=\alpha t e_1$,
noting that thanks to
\cref{one} and \ref{wassump}, $A x_0 = b_0$ and that thanks to
\cref{ass:n} \eqref{ass:n1}, $\N = \N + \alpha t e_1$.

@@

From these results, by~\cref{maxnorm} it immediately follows that $\hg$ is an $\alpha$-contraction with respect to the $\max$-norm:

@@

The operator $\hg \colon \R^n\ra \R^n$ obeys the $\max$-norm contraction property with factor $\alpha$.
\if0
 and the following iterative scheme based on the ALUB projection operator $\hg$, see \eqref{apvi}, converges to a unique fixed point $\hv$.

@@

\fi

@@

As noted, this is immediate from \cref{tgmonotone,tgshift} and \cref{maxnorm}.
%Follows along similar lines as the proof of Theorem~\ref{gmaxcontra}.

@@

In what follows we denote by $\hv$ the unique fixed point of $\hg$:
\begin{align*}
\hv = \hg \hv\,.
\end{align*}
Recall that $\hj$ denotes the solution $\hj$ of the GRLP \eqref{grlp}. 
The vector $\hj$ dominates the vector $\hv$:

@@

The vectors $\hv,\hj$ obey $\hj\geq\hv$.

@@

I think the proof would be much easier to understand if we introduced $r_{c,J} = \argmin_{r} \{ c^\top \Phi r \,:\, W^\top E \Phi r \ge W^\top H J\}$. We could start by saying we need $V_1\le V_0 \le \hj$ because then Lemma~\ref{tgmonotone} will give the desired result. Next, $V_0\le \hj$ holds because on the one hand, $\hj(i) \ge (\Phi r_i)(i)$ for any $i$, while on the other hand,

@@

The next result bounds the distance between $J^*$ and $\hv$ in terms of the distance of $\bj = \Gamma J^*$ and $J^*$, and that between $\bj$ and $\hg J^*$.

@@

We have

@@

||J^*-\hv||_\infty\leq \frac{||J^*-\Gamma J^*||_\infty+||\Gamma J^*-\hg J^*||_\infty}{1-\alpha}.

@@

By the triangle inequality,

@@

%&=\frac{\epsilon+\beta}{1-\alpha}

@@

and so by reordering and another triangle inequality,
\begin{align*}
\norm{J^*-\hv}_{\infty} \le \frac{\norm{ J^*-\hg J^*}_{\infty}}{1-\alpha}
\le \frac{\norm{ J^*-\Gamma J^*}_{\infty}+\norm{\Gamma J^* - \hg J^*}_{\infty}}{1-\alpha}\,.
\end{align*}

@@

\noindent Combining this result with \cref{bestbnd} gives the following corollary:

@@

%Let $\hv$, $\bj$ be as in Theorem~\ref{mt1} and let $r^*\eqdef\argmin_{r\in \R^k}||J^*-\Phi r||_\infty$, then
We have

@@

%\begin{proof}
%The result is obtained by using Lemma~\ref{bestbnd} to replace the term $||J^*-\bj||_\infty$ in Theorem~\ref{mt1}.
%\end{proof}

### 2015-06-14 lubp.tex
ALUBP cleaned

%!TEX root =  autocontgrlp.tex
\section{Least Upper Bound Projection}\label{sec:lubp}
Let $e_j$ denote the $j$th basis vector in the standard Euclidean bases of $\R^n$.
We make the following assumption on $\N$:
\begin{assumption}
\label{ass:n}
The set $\N$ satisfies the following conditions:
\begin{enumerate}[(i)]
\item \label{ass:n1}
For any $t\in \R$, $t e_1 + \N \eqdef \{ r + t e_1 \,:\, r\in \N \} = \N$, i.e., 
$\N$ is invariant to shifts by vectors of the form $t e_1$.
\item \label{ass:n2}
For $2\le k \le n$, $\sup_{r\in \N} | e_k^\top r | <+\infty$.
\item \label{ass:n3}
For each $1\le i \le n$, there exists a solution $r_i\in \R^k$ to the ALP \eqref{alp} with $c = e_i$ such that
$r_i\in \N$.
\end{assumption}

@@

\F_J\eqdef\{\,\Phi r\,|\,\Phi r\geq TJ, r\in \N\,\}\,.

@@

Further, since $(\Gamma J)(i) = \min\{ V(i) \,|\, V\in \F_J \}$, it also follows that $\Gamma J \le V$ holds for any $V\in \F_J$.

@@

We now show that $\Gamma$ is a contraction. 
By \cref{maxnorm}, it suffices to show that $\Gamma$ is monotone and linear along $[\one]$-rays.
We start with monotonicity.
\begin{lemma}\label{gmonotone}
For $J_1, J_2\in \R^n$ such that $J_1\geq J_2$, we have $\Gamma J_1\geq \Gamma J_2$.
\end{lemma}
\begin{proof}
Choose any $i\in \{1,\ldots,n\}$. 
Since $J_1\geq J_2$, we have $TJ_1\geq TJ_2$.
Hence, $\F_{J_1} \subset \F_{J_2}$ and thus $(\Gamma J_1)(i) \ge (\Gamma J_2)(i)$. 
Since $i$ was arbitrary, the result follows.
\end{proof}
\begin{lemma}\label{lpsol}
Let $A\in \R^{u\times v}$, $b,c\in \R^u$ and $b_0=Ax_0$ for 
some $x_0 \in \R^v$, $\N \subset \R^v$ such that $\N =x_0+ \N$. Then
\begin{align}
\min\{c^\top Ax:Ax\geq b+b_0, x\in \N\} =\min\{c^\top A y:Ay \geq b, y \in \N \}+c^\top b_0.
\end{align}
\end{lemma}
\begin{proof}
The claim follows by the change of variables $y := x-x_0$.
\end{proof}
\begin{lemma}\label{gshift}
Let $J\in \R^n$ and $t\in \R$ be a constant.
Then, $\Gamma (J+t\one) = \Gamma J + \alpha t \one$.
\end{lemma}
\begin{proof}
Consider the $i^{th}$ linear programs associated with $\Gamma J$ and $\Gamma (J+ t\one)$. 
The result follows by using Lemma~\ref{lpsol} with $A=\Phi$, $b=TJ$, $c=e_i$, $b_0=\alpha t\mathbf{1}$ 
and $x_0=\alpha t e_1$, noting that by our assumption on $\N$, $\N = \N + \alpha t e_1$.
\end{proof}
\begin{theorem}\label{gmaxcontra}
The operator $\Gamma  \colon \R^n\ra \R^n$ obeys the $\max$-norm contraction property with factor $\alpha$.
\end{theorem}
\begin{proof}
The statement follows immediately from Lemmas~\ref{gmonotone} and~\ref{gshift}, and \cref{maxnorm}.
\end{proof}
\if0
\begin{corollary}
Let $V_0$ be arbitrary and $(V_n)$ be defined recursively by
\begin{align}\label{pvi}
V_{n+1}&=\Gamma V_n, \quad n\geq 0\,.
\end{align}
Then, $V_n$ converges to the the unique fixed point $\tv$ of $\Gamma$.
\end{corollary}
\fi
Define $\tv$ denote the unique fixed point of $\Gamma$.
\todoc{Poor choice of symbols: $\tv$ is the fixed point of $\Gamma$ and $\hat{V}$ will be the fixed point of $\tilde{\Gamma}$? Hmm.. Not a high priority to change it, but still..}
\begin{lemma}\label{gfp}
The vector $\tv$ obeys $\tv\geq T\tv$.
\end{lemma}
\begin{proof}
From the definitions we have $\tv(i) = (\Gamma \tv)(i) = (\Phi r_{e_i,\tv})(i) \ge (T\tv)(i)$, $1\le i \le n$.
\if0
Consider the $i^{th}$ linear program associated with $\Gamma \tv$. We know that $\Phi r_{e_i}\geq T \tv$, for all $i=1,\ldots, n$. The result follows from noting that $\tv$ is the unique fixed point of $\Gamma $ and that $\tv(i)=(\Phi r_{e_i,V})(i)$.
\fi
\end{proof}
Recall that $\tj$ denotes the solution to the ALP in \eqref{alp}.
The fixed point of $\Gamma$ is sandwiched between $J^*$ and $\tj$:
\begin{lemma}\label{relation1}
%Assume that $\N$ is such that if $r_i$ is
%the solution to the ALP in \eqref{alp} for $c=e_i$ then $r_i\in \N$, $1\le i \le n$.
% $\tv$, the unique fixed point of the iterative scheme \eqref{pvi}, and the solution $\tj$ to the ALP in \eqref{alp}, obey 
The relation $\tj\geq\tv\geq J^*$ holds.
\end{lemma}
\begin{proof}
Since $\tv\geq T\tv$ it follows that $\tv\geq J^*$. 
Let $r_1, r_2,\ldots, r_n$ be solutions to the ALP in \eqref{alp} for $c=e_1, e_2,\ldots,e_n$ respectively
such that $r_1,\dots,r_n\in \N$. It is possible to find such solutions thanks to \cref{ass:n} \eqref{ass:n3}.
Define $V_0\in \R^n$ by $V_0(i)=\underset{j=1,\ldots, n}{\min}(\Phi r_j)(i)$, $1\le i \le n$. 
It is clear from the definition of $V_0$ that $\tj(i), \Phi r_j(i) \geq\Phi r_i(i)\geq V_0(i)$, $1\le i,j \le n$. 
Also from the monotone property of $T$, we have $T\Phi r_j\geq T V_0$,
hence we also have,
$\Phi r_j\geq T\Phi r_j \geq T V_0$, 
and so by taking component-wise minimum,
$V_0 \geq T V_0$.

>>>>

%!TEX root =  autocontgrlp.tex
\section{Least Upper Bound Projection}\label{sec:lubp}
Let $e_j$ denote the $j$th basis vector in the standard Euclidean bases of $\R^n$.
We make the following assumption on $\N$:
\begin{assumption}
\label{ass:n}
The set $\N$ satisfies the following conditions:
\begin{enumerate}[(i)]
\item \label{ass:n1}
For any $t\in \R$, $t e_1 + \N \eqdef \{ r + t e_1 \,:\, r\in \N \} = \N$, i.e., 
$\N$ is invariant to shifts by vectors of the form $t e_1$.
\item \label{ass:n2}
For $2\le i \le k$, $\sup_{r\in \N} | e_i^\top r | <+\infty$.
\item \label{ass:n3}
For each $1\le i \le n$, there exists a solution $r_i\in \R^k$ to the ALP \eqref{alp} with $c = e_i$ such that
$r_i\in \N$.
\end{enumerate}
\end{assumption}
Recalling that the first column vector of $\Phi$ is $\one$, \eqref{ass:n1} implies that $\{\Phi r \,:\,r\in \N \}$ is invariant to shifts by constants, a property that will prove to be useful.
Part~\eqref{ass:n2} is required to guarantee that solutions to the LPs we define below are bounded.
To satisfy \eqref{ass:n2} and \eqref{ass:n3} simultaneously, it is enough to select $\N$ to be contained in 
$\{ r = (r_1,\dots,r_k)^\top \in \R^k \,:\, \max_{2\le i \le k} |r_i| \le r_{\max} \}$ with $r_{\max}>0$ sufficiently large.
\todoc{Can we calculate an a priori bound on how $r_{\max}$ should be? It is a major weakness if we cannot
suggest a way to define $\N$.}

@@

\F_J\eqdef\{\,\Phi r\,:\,\Phi r\geq TJ, r\in \N\,\}\,.

@@

Further, since $(\Gamma J)(i) = \min\{ V(i) \,:\, V\in \F_J \}$, it also follows that $\Gamma J \le V$ holds for any $V\in \F_J$.

@@

We now show that $\Gamma$ is a contraction. 
By \cref{maxnorm}, it suffices to show that $\Gamma$ is monotone and linear along $[\one]$-rays with a factor smaller than one.
We start with monotonicity.
\begin{lemma}\label{gmonotone}
For $J_1, J_2\in \R^n$ such that $J_1\leq J_2$, we have $\Gamma J_1\leq \Gamma J_2$.
\end{lemma}
\begin{proof}
Choose any $i\in \{1,\ldots,n\}$. 
Since $J_1\leq J_2$, we have $TJ_1\leq TJ_2$.
Hence, $\F_{J_2} \subset \F_{J_1}$ and thus $(\Gamma J_1)(i) \le (\Gamma J_2)(i)$. 
Since $i$ was arbitrary, the result follows.
\end{proof}
\begin{lemma}\label{lpsol}
Let $A\in \R^{u\times v}$, $b,c\in \R^u$ and $b_0=Ax_0$ for 
some $x_0 \in \R^v$, $\N \subset \R^v$ such that $\N =x_0+ \N$. Then
\begin{align}
\min\{c^\top Ax:Ax\geq b+b_0, x\in \N\} =\min\{c^\top A y:Ay \geq b, y \in \N \}+c^\top b_0.
\end{align}
\end{lemma}
\begin{proof}
The claim follows by the change of variables $y := x-x_0$.
\end{proof}
\begin{lemma}\label{gshift}
Let $J\in \R^n$ and $t\in \R$ be a constant.
Then, $\Gamma (J+t\one) = \Gamma J + \alpha t \one$.
\end{lemma}
\begin{proof}
Consider the $i^{th}$ linear programs associated with $\Gamma J$ and $\Gamma (J+ t\one)$. 
The result follows by using Lemma~\ref{lpsol} with $A=\Phi$, $b=TJ$, $c=e_i$, $b_0=\alpha t\mathbf{1}$ 
and $x_0=\alpha t e_1$, noting that thanks to
\cref{one}, $A x_0 = b_0$ and that thanks to
\cref{ass:n} \eqref{ass:n1}, $\N = \N + \alpha t e_1$.
\end{proof}
\begin{theorem}\label{gmaxcontra}
The operator $\Gamma  \colon \R^n\ra \R^n$ obeys the $\max$-norm contraction property with factor $\alpha$.
\end{theorem}
\begin{proof}
The statement follows immediately from Lemmas~\ref{gmonotone} and~\ref{gshift}, and \cref{maxnorm}.
\end{proof}
\if0
\begin{corollary}
Let $V_0$ be arbitrary and $(V_n)$ be defined recursively by
\begin{align}\label{pvi}
V_{n+1}&=\Gamma V_n, \quad n\geq 0\,.
\end{align}
Then, $V_n$ converges to the the unique fixed point $\tv$ of $\Gamma$.
\end{corollary}
\fi
Define $\tv$ denote the unique fixed point of $\Gamma$.
\todoc{Poor choice of symbols: $\tv$ is the fixed point of $\Gamma$ and $\hat{V}$ will be the fixed point of $\tilde{\Gamma}$? Hmm.. Not a high priority to change it, but still..}
\begin{lemma}\label{gfp}
The vector $\tv$ obeys $\tv\geq T\tv$.
\end{lemma}
\begin{proof}
From the definitions we have $\tv(i) = (\Gamma \tv)(i) = (\Phi r_{e_i,\tv})(i) \ge (T\tv)(i)$, $1\le i \le n$.
\if0
Consider the $i^{th}$ linear program associated with $\Gamma \tv$. We know that $\Phi r_{e_i}\geq T \tv$, for all $i=1,\ldots, n$. The result follows from noting that $\tv$ is the unique fixed point of $\Gamma $ and that $\tv(i)=(\Phi r_{e_i,V})(i)$.
\fi
\end{proof}
Recall that $\tj$ denotes the solution to the ALP in \eqref{alp}.
The fixed point of $\Gamma$ is sandwiched between $J^*$ and $\tj$:
\begin{lemma}\label{relation1}
%Assume that $\N$ is such that if $r_i$ is
%the solution to the ALP in \eqref{alp} for $c=e_i$ then $r_i\in \N$, $1\le i \le n$.
% $\tv$, the unique fixed point of the iterative scheme \eqref{pvi}, and the solution $\tj$ to the ALP in \eqref{alp}, obey 
The relation $\tj\geq\tv\geq J^*$ holds.
\end{lemma}
\begin{proof}
Since $\tv\geq T\tv$ it follows that $\tv\geq J^*$. 
Let $r_1, r_2,\ldots, r_n$ be solutions to the ALP in \eqref{alp} for $c=e_1, e_2,\ldots,e_n$ respectively
such that $r_1,\dots,r_n\in \N$. It is possible to find such solutions thanks to \cref{ass:n} \eqref{ass:n3}.
Define $V_0\in \R^n$ by $V_0(i)=\underset{j=1,\ldots, n}{\min}(\Phi r_j)(i)$, $1\le i \le n$. 
It is clear from the definition of $V_0$ that $\tj(i), \Phi r_j(i) \geq\Phi r_i(i)\geq V_0(i)$, $1\le i,j \le n$. 
Also from the monotone property of $T$, we have $T\Phi r_j\geq T V_0$,
hence we also have,
$\Phi r_j\geq T\Phi r_j \geq T V_0$, 
and so by taking component-wise minimum,
$V_0 \geq T V_0$.
